---
title: "R - Quiz"
author: "null"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---


### Question 1 

What function should I use to load a csv file into a data frame?

```{r}
read.csv()
```

1. read.csv() 
2. readCsv() 
3. read_CSV() 
4. csv.read() 


### Question 2

Select a library for data transformation (data wrangling).

```{r, include=FALSE, echo=FALSE, warning=FALSE}
dplyr::
tidyr::
```

1. ggplot2
2. plotly
3. dplyr
4. xlsx

### Question 3 

What will be the result of calling the following function?

```{r message=FALSE, warning=FALSE, eval=TRUE}

SampleFunction <- function(a, b) {
 a <- 1
 c <- a + b
 
 return(c)

}

SampleFunction(a= 2, b = 3)

```


1. 4
2. 5
3. 6
4. 7

### Question 4 

Which of the following libraries is not used to create graphs?

1. ggplot2
2. plotly
3. dplyr
4. None of the above.

### Question 5

What will be the result of executing the following line of code?

```{r message=FALSE, warning=FALSE, eval=FALSE}
x <- c("a", "b", "c", "d")
y <- c("w", "x", "y", "z")


paste(x, y, sep="%%")
```

1. "a;w","b;x","c;y","d;z"
2. "a%%w%%b%%x%%c%%y%%d%%z"
3. "a,w,b,x,c,y,d,z"
4. "a%%w" "b%%x" "c%%y" "d%%z"


### Question 6 

What code should be entered to select a record from the 2nd row and 3rd column of the data frame below?

```{r message=FALSE, warning=FALSE, eval=FALSE}

SampleDF <- data.frame(City = c("Gdansk", "Gdynia", "Sopot", "Gdansk", "Gdynia", "Sopot", "Sopot", "Gdansk"),
                  Sales = c(500, 400, 200, 400, 200, 250, 200, 100 ),
                  Employees = c(6, 4, 1, 6, 4, 1, 2, 3 ))

SampleDF[2,3]

```


1. SampleDF[,2]
2. Df1[2,3]
3. SampleDF[2,3]
4. SampleDF[3,2]

### Question 7 

What type of graph will the following code produce?

```{r message=FALSE, warning=FALSE, eval=FALSE}
library(ggplot2)
library(datasets)

stocks <- data.frame(EuStockMarkets)

plot1<- ggplot(data = stocks,
                        mapping = aes( x = FTSE,
                                       y = DAX))

plot2 <- plot1 + geom_point()
```

1. Scatter plot
2. Line plot
3. Histogram
4. Map

### Question 8 

What type(s) of data will/will be stored in the vector below?

```{r}
SampleVec <- c(1, "bla", TRUE)
class(SampleVec)
```

1. numeric, character, logical
2. character
3. numeric, character
4. numeric

### Question 9

What is the '$' character used for in data frames?
```{r}
SampleDF$City
```

1. Referring to individual lines.
2. Build functions.
3. Building relationships between the lines.
4. Referring to individual columns.

### Question 10 

Which of the following lines of code will select the record(s) where sales are greater than 200 and less than 400. 

```{r message=FALSE, warning=FALSE, eval=FALSE}

SampleDF <- data.frame(City = c("Gdansk", "Gdynia", "Sopot", "Gdansk", "Gdynia", "Sopot", "Sopot", "Gdansk"),
                  Sales = c(500, 400, 200, 400, 200, 250, 200, 100 ),
                  Employees = c(6, 4, 1, 6, 4, 1, 2, 3 ))
SampleDF %>% filter(Sales > 200 & Sales < 400)
```


1. SampleDF %>% filter(Sales > 200 & Sales < 400)
2. SampleDF %>% filter(Sales > 200 | Sales < 400)
3. SampleDF %>% filter(Sales >= 200 & Sales =< 400)
4. SampleDF %>% filter(Sales > 200 %and% Sales < 400)


### Question 11 

For the vector below, specify which code will not only call "MO" and "TU".

```{r message=FALSE, warning=FALSE, eval=FALSE}
t <- c("MO", "TU", "WE", "TH", "FR", "SA", "SU")

t[c(1,2)]
t[c(-3,-4,-5,-6,-7)]
t[t=="MO" | t=="TU"]

t[c(-7:-5)]
```


1. t[c(1,2)] 
2. t[c(-3,-4,-5,-6,-7)] 
3. t[t=="MO" | t=="TU"]
4. t[c(-7:-5)]


### Question 12 

If the following columns are available in the data frame:
"name", "year", "month", "day", "hour"

What will the code below trigger?

```{r message=FALSE, warning=FALSE, eval=FALSE}
library(datasets)

data(storms)
storms %>% mutate(Date = paste(year, month, day, sep= "-")) 

```

1. The lines that are in the format "yyyy-mm-dd" will be selected.
2. It will replace the current "Date" column, which will consist of the year, month and day.
3. It will create a new "Date" column, which will consist of the year, month and day.
4. None of the above.

### Question 13

What will be the result of running the following lines of code?

```{r}
library(dplyr)

day = c(1, 1, 2, 2, 3, 3)
hour = c(8, 16, 8, 16, 8, 16)
profit = c(100, 200, 50, 60, NA, NA)
shop.data = data.frame(day, hour, profit)


shop.data %>% mutate(profit= ifelse(is.na(profit), median(profit, na.rm=TRUE), profit))


```

1. All records in the profit column will be converted to NA.
2. NA records will be replaced with the arithmetic mean of the entire profit column.
3. NA records will be replaced with the median value of the entire profit column.
4. Records that are not NA will be replaced with the median value of the entire profit column.

### Question 14

What will the following lines of code do?

```{r message=FALSE, warning=FALSE, eval=FALSE}
library(ggplot2)
library(datasets)

stocks <- data.frame(EuStockMarkets)

plot1 <- ggplot(data = stocks, 
                aes(DAX, FTSE ))


plot1
```


1. A complete chart will appear in the plot tab (background with scale, line chart and signed axes).
2. An outline of the chart will appear in the plots tab (background with a scale and labeled axes).
3. A complete graph will appear in the plot tab (background with scale, scatter plot and signed axes).
4. Nothing will be displayed.


### Question 15

What will appear on the screen (in the console) after running the following line of code:

```{r message=FALSE, warning=FALSE, eval=FALSE}

for ( i in  10 : 1) {
  print(i)
}



```


1. Vector filled with values from 1 to 10.
2. Vector filled with values from 10 to 1.
3. 10 individual elements (single element vectors) with values from 1 to 10.
4. 10 individual elements (one-element vectors) with values from 10 to 1.


### Question 16

What function should be inserted in "__" to obtain the "strong wind" wind category in the "Cat1" column when the speed exceeds 20 (km/h).

```{r message=FALSE, warning=FALSE, eval=FALSE}


tst2 <- storms %>%
  mutate(Cat1 = ifelse(wind > 20, "Strong Wind", "Weak Wind"))



```

1. if
2. ifelse
3. countif
4. elseif


### Question 17

What functions/steps will you use to build summaries (aggregations) for individual cities from the data frame below?


```{r}
library(dplyr)
SampleDF <- data.frame(City = c("Gdansk", "Gdynia", "Sopot", "Gdansk", "Gdynia", "Sopot", "Sopot", "Gdansk"),
                  Sales = c(500, 400, 200, 400, 200, 250, 200, 100 ),
                  Employees = c(6, 4, 1, 6, 4, 1, 2, 3 ))

SampleDF %>% group_by(City) %>% summarise(Sum = sum(Sales), Average = mean(Sales))

```

1. SampleDF %>% group_by(City) %>% summarise(Sum = sum(Sales), Average = mean(Sales))
2. SampleDF %>% summarise(Sum = sum(Sales), Average = mean(Sales)) %>% group_by(City)
3. TestDF %>% group_by(City) %>% summarise(Sum = sum(Sales), Average = mean(Sales))
4. TestDF %>% summarise(Sum = sum(Sales), Average = mean(Sales))

### Question 18

Which of the following codes will return an error?

```{r}
mtcars %>% filter(cyl == 6 && disp >= 160)
mtcars %>% filter(cyl == 6 || disp >= 160)
mtcars %>% filter(cyl = 6 | disp >= 160)
mtcars %>% filter(cyl == 6 | disp >= 160)
```

1. mtcars %>% filter(cyl == 6 && disp >= 160)
2. mtcars %>% filter(cyl == 6 || disp >= 160)
3. mtcars %>% filter(cyl = 6 | disp >= 160)
4. mtcars %>% filter(cyl == 6 | disp >= 160)

### Question 19

Which command will create a box plot?

1. geom_box()
2. geom_plot()
3. geom_boxplot()
4. geom_jitter()

### Question 20

What will be the result of the following logical operation?

(5 > 3) & !(3 == 4)
```{r}
(5 > 3) & !(3 == 4)
```


1. TRUE
2. FALSE
3. vector (TRUE, TRUE)
4. vector (TRUE, FALSE)

