---
title: "Data Analysis with R language"
output:
  html_document:
    df_print: paged
---

Table of contents:

1. Explanatory data analysis with libraries *dplyr*, *tidyr* and *stringr* \
  * Subsetting columns \
  * Subsetting rows \
  * Logical operators, Boole'an algebra, de Morgan laws \
  * Creating new columns (1 exercise) \
  * Handling missing values \
  * Text manipulation (3 exercises) \
  * Data aggregation (1 exercise) \
  * Pivot tables, long and wide data format \
  * Joining tables \
  
2. Data vizualization with library *ggplot2* \
  * single numerical variable \
  * single categorical variable \
  * pair of numerical variables \
  * pair of categorical variables \
  * numerical and categorical variable \
  * bubble charts (3 exercises) \
  * multiple plot on single panel (2 exercises).

Useful links: \
 - [dplyr cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf) \
 - [tidyr cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/tidyr.pdf) \
 - [stringr cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/strings.pdf) \
 - [ggplot2 cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf) \
 - [A. Kassambara - Guide to Create Beautiful Graphics in R](http://www.sthda.com/english/download/3-ebooks/5-guide-to-create-beautiful-graphics-in-r-book/).
 
[Color palletes](https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/)
  
```{r}
if (!require('tidyverse')) install.packages('tidyverse')
library(tidyverse)
```

Dataset comes from [https://flixgem.com/](https://flixgem.com/) (version from 12th March 2021). Dataset includes info about 9425 movies and series available on Netflix.
```{r}
df = read.csv('netflix-dataset.csv')
```

```{r}
# top_n(groupby(filter(select(df, ...))))
# 
# df %>%
#   select() %>%
#   filter() %>%
#   group_by() %>%
#   top_n()
```

## Exploratory data analsis with dplyr and tidyr

### Subsetting columns

We select column by their names using *select()*. We can also remove columns be preceeding its name with minus sign.
```{r}
df %>%
  select(Title, Runtime, IMDb.Score, Release.Date)

df2 = df %>%
  select(-Title, -Runtime)

df %>%
  select(-Title, -Runtime)

df %>%
  select(Title:Runtime)

df %>%
  select(1, 5, 8)
```

Useful functions during selecting columns: \
 - *starts_with()* - select/remove columns which names begins with a given string \
 - *ends_with()* - select/remove columns which names end with a given string \
 - *contains()* - select/remove columns which names contains given string.
```{r}
df %>%
  select(ends_with('score'))

df %>%
  select(starts_with('IMDb', ignore.case = FALSE))
```

With function *matches()* we can select or remove columns which names include given regular expression. Useful tool for building and testing regular expressions can be found here: [https://regex101.com/](https://regex101.com/).

```{r}
df %>%
  select(matches('^[A-Za-z]{5,6}$'))
```

Function *select()* always returns data frame, but sometimes we need a single column returned as a vector. It can be achieved with function *pull()*.
```{r}
df3 = df %>% pull(Title)

df %>%
  pull(IMDb.Score, Title)
```

### Filtering rows

We filter rows with function *filter()* and comparison operators *==*, *!=*, *>*, *>=*, *<*, *<=*, *between()*. 
```{r}
unique(df$Series.or.Movie)
table(df$Series.or.Movie)
```

```{r}
df %>%
  filter(Series.or.Movie == 'Series')

df %>%
  filter(IMDb.Score > 8.5)

df %>%
  filter(IMDb.Score >= 7 & IMDb.Score <= 9)

df %>%
  filter(between(IMDb.Score, 7, 9))
```

### Logical operators, Boole'an algebra, de Morgan laws

Logical operator *AND* we denote with symbol *&* \
 - *FALSE & FALSE = FALSE* \
 - *FALSE & TRUE = FALSE* \
 - *TRUE & FALSE = FALSE* \
 - *TRUE & TRUE = TRUE*
```{r}
df %>%
  filter(IMDb.Score > 8.5 & Series.or.Movie == 'Series')
```

Logical operator *OR* we denote with symbol *|* \
 - *FALSE | FALSE = FALSE* \
 - *FALSE | TRUE = TRUE* \
 - *TRUE | FALSE = TRUE* \
 - *TRUE | TRUE = TRUE*
```{r}
df %>%
  filter(IMDb.Score > 8.5 | Series.or.Movie == 'Series')
```

De Morgan laws states then when entering with negation into bracket then OR and AND are being exchanged: \
*not (A & B) = (not A) |(not B)* \
*not (A | B) = (not A) & (not B)* \
```{r}
df %>%
  filter(!(IMDb.Score > 8.5 & Series.or.Movie == 'Series'))

df %>%
  filter(!(IMDb.Score > 8.5) | !(Series.or.Movie == 'Series'))

df %>%
  filter((IMDb.Score <= 8.5 | Series.or.Movie != 'Series'))
```

### Creating new columns:

With function *mutate()* we can add new columns or overwrite existing ones.
```{r}
df %>%
  mutate(
    Score_category = ifelse(IMDb.Score >= 3, 'Good', 'Poor')
    , Popularity = ifelse(IMDb.Votes > 1e6, 'High', 'Low')
    , .before = 1
  )

df = df %>%
  mutate(
    Release.Date = as.Date(Release.Date, format = '%m/%d/%Y')
    ,  Netflix.Release.Date = as.Date(Netflix.Release.Date, format = '%m/%d/%Y')
  )
```

**EXERCISE 1:** What is the oldest movie directed by Woody Allen available on Netflix?
```{r}
# df2 = df %>%
#   mutate(
#   The_oldest = ifelse(Release.Date > max(Release.Date)-1 & Director=="Woody Allen", 'Woddy', 'Nie')
# )

df %>%
  filter(Director == 'Woody Allen')

df %>%
  filter(Director == 'Woody Allen') %>%
  filter(Release.Date == min(Release.Date)) %>%
  select(Title, Director, Release.Date)

# df %>%
#   group_by(Director) %>%
#   filter(Release.Date == min(Release.Date)) %>%
#   select(Title, Director, Release.Date)
```

When using *case_when()* we don't need to explicitly write mutually exclusive conditions. Function evaluation occurs just after it encounters first true set of conditions.
```{r}
df %>%
  mutate(Score_category = case_when(
    IMDb.Score <= 2 ~ 'Very Poor'
    , IMDb.Score <= 4 ~ 'Poor'
    , IMDb.Score <= 6 ~ 'Medium'
    , IMDb.Score <= 8 ~ 'Good'
    , IMDb.Score <= 10 ~ 'Very Good'
  ))
```

Mathematical functions performed on each row we apply by precedding whole operation with *rowwise()* function.
```{r}
# mean(c(2,4, NA), na.rm = TRUE)
df %>%
  rowwise() %>%
  mutate(Avg_score = mean(
    c(IMDb.Score * 10, Hidden.Gem.Score * 10, Rotten.Tomatoes.Score, Metacritic.Score), na.rm = TRUE
  ) %>% round(1)) %>%
  select(Title, Avg_score)
```

By default, new columns created with *mutate()* are added at the end of the data frame. With *relocate()* we can change position of a particular column.
```{r}
df %>%
  relocate(Director, .before = Title)
```

We can rename columns with *rename()*.
```{r}
df %>%
  rename(
    Title_of_movie = Title
    , Genre_of_movie = Genre
  )
```

### Handling missing values

Library *tidyr* provides several handful of function for dealing with missing data: \
 - *drop_na()* - delete rows with missing values in selected columns \
 - *replace_na()* - replace missing values with constant value \
 - *fill()* - replace missing values with first or last non-missing value.

```{r}
# Check missing values
sum(is.na(df$Hidden.Gem.Score))
sapply(df, function(x) sum(is.na(x)))

df %>%
  na.omit()

df %>%
  drop_na(Title, Director)

# Replace with average value
df2 = df %>%
  mutate(Hidden.Gem.Score = replace_na(Hidden.Gem.Score, mean(Hidden.Gem.Score, na.rm = TRUE)))

sum(is.na(df2$Hidden.Gem.Score))

fill()
```

### Text manipulation

Library *stringr* contains handful of useful function for string manipulation and regular expressions. Most of functions from this library beging with *str_*.

```{r}
str_
```

Question: What can be improved in the below code to maintain *tidyverse* syntax?
```{r}
genres_df = df %>%
  pull(Genre) %>%
  str_c(collapse = ', ') %>%
  str_extract_all('[A-Za-z\\-]+', simplify = TRUE) %>%
  table() %>%
  as.data.frame() %>%
  arrange(-Freq)
```

```{r}
df %>%
  mutate(Poland_available = str_detect(Country.Availability, 'Poland'), .before = 1) %>%
  filter(Poland_available == TRUE)

df %>%
  filter(str_detect(Country.Availability, 'Poland'))

# str_replace_all()
# str_re
```


With function *separate()* we can split single column into several ones, and we can join multiple columns into a single one with function *unite()*.
```{r}
df2 = df %>%
  unite(
    col = 'Scores'
    , Hidden.Gem.Score, IMDb.Score, Rotten.Tomatoes.Score, Metacritic.Score
    , sep = ', '
  ) %>% 
  select(Title, Scores)

df2 %>%
  separate(
    col = Scores
    , into = paste('Score', 1:4)
    , sep = ', '
  )
``` 

**EXERCISE 2:** What are the best three comedies available in Polish language?
```{r}
df %>%
  filter(str_detect(Languages,'Polish')) %>%
  filter(str_detect(Genre,'Comedy')) %>%
  select(Title, Hidden.Gem.Score, Genre) %>%
  arrange(Hidden.Gem.Score)%>%
  # top_n(3)
  slice_max(Hidden.Gem.Score, n = 3)
```

**EXERCISE 3:** What are the most popular tags for series and movies available in Polish language?
```{r}
df %>%
  filter(str_detect(Languages, 'Polish')) %>%
  pull(Tags) %>%
  str_c(collapse= ',') %>%
  # str_extract_all('[A-Za-z\\-\\s\\&]+', simplify = TRUE) %>%
  str_split(',', simplify = TRUE) %>%
  table() %>%
  as.data.frame()%>%
  arrange(-Freq)
```

### Data aggregation

With functions *group_by()* and *summarize()* we work on aggregated data.
```{r}
df %>%
  group_by(Director) %>%
  filter(Release.Date == min(Release.Date, na.rm = TRUE)) %>%
  select(Title, Director, Release.Date)

df %>%
  group_by(Series.or.Movie) %>%
  summarise(
    Count = n()
    , Avg_score = mean(IMDb.Score, na.rm = TRUE) %>% round(2)
    , Avg_votes = mean(IMDb.Votes, na.rm = TRUE) %>% round(2)
    , Awards_total = sum(Awards.Received, na.rm = TRUE)
  )

df %>%
  group_by(Series.or.Movie, Runtime) %>%
  summarise(
    Count = n()
    , Avg_score = mean(IMDb.Score, na.rm = TRUE) %>% round(2)
    , Avg_votes = mean(IMDb.Votes, na.rm = TRUE) %>% round(2)
    , Awards_total = sum(Awards.Received, na.rm = TRUE)
  )

```

```{r}
library(lubridate)
month(df$Release.Date[1:5])

# lubridate::year(df$Release.Date)
```

**EXERCISE 4:** For movies and series released in 2019 and 2020 what is the average time between its release date and appearance on Netflix?

1. extract year from the date
2. filter 2019 and 2020
3. calcualte difference between appearance on Netflix and release
4. (two-step operation) calculate averege difference per 2019 and 2020
```{r}
df %>%
  mutate(Year = lubridate::year(Release.Date), .before = 1) %>%
  filter(Year == 2019 | Year == 2020) %>%
  mutate(Days = Netflix.Release.Date - Release.Date, .before = 1) %>%
  group_by(Year) %>%
  summarise(Avg = mean(Days, na.rm = TRUE) %>% round(0))
```

**EXERCISE 5:** What are average ratings of movies produced in different decades (i.e. 60s, 70s, 80s, 90s etc.)?
```{r}
# 3rd digit of year is a good indicator of a decade
df %>%
  mutate(Year = lubridate::year(Release.Date), .before = 1) %>%
  mutate(Decade = str_c(str_sub(Year, 3, 3), '0'), .before = 1) %>%
  group_by(Decade) %>%
  summarise(Count = n(), Avg = mean(IMDb.Score, na.rm = TRUE) %>% round(2)) %>%
  arrange(-Avg)

```

### Pivot tables, long and wide data format

df in *wide* format: \
 - single individuals are represented by rows \
 - their attributes are represented by columns \
 - in each cell there is a value of a given attribute for a given individual.
 
df in *long* format: \
 - in first columns can can see individual key (which can be constructed from multiple columns) \
 - all the attributes are in single column \
 - all the values are in single column.

*Long* format is especially useful when creating plots with library *ggplot2*.

```{r}
df2 = df %>%
  select(Title, ends_with('Score'))
```


```{r}
df3 = df2 %>%
  pivot_longer(-Title)

# df3 = df2 %>%
#   pivot_longer(
#     cols = 2:5
#     , names_to = 'Portal'
#     , values_to = 'Score'
#   )


df4 = df3 %>%
  pivot_wider(
    id_cols = 1
    , names_from = 'name'
    , values_from = 'value'
  )
```


### Joining tables

```{r}
metacritic_ratings = df %>%
  select(Title, Metacritic.Score) %>%
  .[1:100, ] %>%
  drop_na()

rotten_tomatoes_ratings = df %>%
  select(Title, Rotten.Tomatoes.Score) %>%
  .[1:100, ] %>%
  drop_na()

imdb_ratings = df %>%
  select(Title, IMDb.Score) %>%
  .[1:100, ] %>%
  drop_na()
```

We join tables by their respective keys, exactly how we do it in SQL.
```{r}
left_join()
right_join()
inner_join()
full_join()
anti_join()

anti_join(
  x = metacritic_ratings
  , y = rotten_tomatoes_ratings
  # , by = 'Title'
  , by = c('Title' = 'Title')
)

metacritic_ratings %>%
  left_join(
    y = rotten_tomatoes_ratings
    , by = 'Title'
  ) %>%
  left_join(
    y = imdb_ratings
    , by = 'Title'
  )
```

## Data vizualization with library ggplot2

```{r}
theme_set(theme_bw())
```

General rules for creating plots with *ggplot2*: \
 - as arguments to *ggplot()* function provide data frame and optionally axes *x*, *y* and parameters *shape*, *color*, *fill*, *group*, *size*, *alpha* \
 - add plot layers with functions starting with *geom_* or *stat_* \
 - modify plot by adding legends, titles, markers etc.
 
For quick data vizualization one can use function *qplot()*, which automatically picks plot type to the underlying data.

```{r}
qplot(df$IMDb.Score)
```

### Single numerical variable

```{r}
ggplot(df, aes(x = IMDb.Score)) +
  geom_histogram(color = 'blue', fill = 'white') +
  labs(x = 'Movie rating', y = 'Count of movies', title = 'Histogram of ratings') +
  theme_bw() -> plot2

ggplot(df, aes(x = IMDb.Score)) +
  geom_histogram(color = 'blue', aes(fill = Series.or.Movie), alpha = 0.5) +
  labs(
    x = 'Movie rating'
    , y = 'Count of movies'
    , fill = 'Serie or movie'
    , title = 'Histogram of ratings'
  ) +
  # theme_bw() +
  # theme(legend.position = 'bottom')
  theme(legend.position = c(0.2, 0.8))
```

### Single categorical variable

```{r}
# install.packages("RColorBrewer")
library(RColorBrewer)
```


```{r}
table(df$Runtime)
df = df %>% filter(Runtime != '')

ggplot(df, aes(y = Runtime)) +
  geom_bar(aes(fill = Runtime)) +
  theme(
    legend.position = 'none'
    , axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  ) +
  scale_y_discrete(
    limits = c(
      '< 30 minutes'
      , '30-60 mins'
      , '1-2 hour'
      , '> 2 hrs'
    )
    , labels = function(x) str_wrap(x, width = 8)
  ) +
  scale_fill_manual(values = brewer.pal(4, 'Pastel1'))
```

### Pair of numerical variables

```{r}
ggplot(df, aes(x = IMDb.Score, y = IMDb.Votes)) +
  geom_point(aes(color = Series.or.Movie)) +
  geom_smooth(aes(color = Series.or.Movie), linewidth = 2)

```

### Pair of categorical variables

```{r}
ggplot(df, aes(x = Series.or.Movie, y = Runtime)) +
  geom_jitter(width = 0.25, height = 0.40, aes(color = Runtime)) +
  theme(legend.position = 'none')
```

### Numerical and categorical variable

```{r}
ggplot(df, aes(x = Series.or.Movie, y = IMDb.Score)) +
  geom_boxplot(aes(fill = Runtime))
```

### Bubble charts
```{r}
# install.packages('ggrepel')
library(ggrepel)
```


```{r}
genres_df = df %>%
  pull(Genre) %>%
  str_c(collapse = ', ') %>%
  str_extract_all('[A-Za-z\\-]+', simplify = TRUE) %>%
  table() %>%
  as.data.frame() %>%
  arrange(-Freq) %>%
  mutate(
    Avg_score = NA
    , Avg_votes = NA
  )

colnames(genres_df)[1] = 'Genre'

for (i in 1:nrow(genres_df)) {
  category = genres_df$Genre[i] %>% as.character()
  
  genres_df_filtered = df %>%
    filter(str_detect(Genre, category))
  
  genres_df$Avg_score[i] = genres_df_filtered %>%
    pull(IMDb.Score) %>%
    mean(na.rm = TRUE)
  
  genres_df$Avg_votes[i] = genres_df_filtered %>%
    pull(IMDb.Votes) %>%
    mean(na.rm = TRUE)
  
}

ggplot(genres_df, aes(x = Avg_score, y = Avg_votes, label = Genre)) +
  geom_point(aes(size = Freq, color = Genre)) +
  geom_label_repel() +
  theme(legend.position = 'none')

```

**EXERCISE 6:** Create alike plot with actors starred in the best movies.
```{r}
actors_df = df %>%
  pull(Actors) %>%
  str_split(', ', simplify = TRUE) %>%
  as.data.frame() %>%
  pivot_longer(cols = 1:4, names_to = 'v1', values_to = 'Actor') %>%
  pull(Actor) %>% 
  table() %>%
  as.data.frame() %>%
  filter(`.` != '') %>%
    mutate(
    Avg_score = NA
    , Avg_votes = NA
  ) %>%
  filter(Freq > 15)

for (i in 1:nrow(actors_df)) {
  actor_name = actors_df$.[i] %>% as.character()
  
  actors_df_filtered = df %>%
    filter(str_detect(Actors, actor_name))
  
  actors_df$Avg_score[i] = actors_df_filtered %>%
    pull(IMDb.Score) %>%
    mean(na.rm = TRUE)
  
  actors_df$Avg_votes[i] = actors_df_filtered %>%
    pull(IMDb.Votes) %>%
    mean(na.rm = TRUE)
  
}

colnames(actors_df)[1] = 'Name'

ggplot(actors_df, aes(x = Avg_score, y = Avg_votes, label = Name)) +
  geom_point(aes(size = Freq, color = Name)) +
  geom_text_repel(max.overlaps = 10) +
  theme(legend.position = 'none')
```

**EXERCISE 7:** Create alike plot to show in which years the best movies were created.

```{r}

```

**EXERCISE 8:** What are the most popular languages available on Netflix? Hist: use column chart.
```{r}

```

### Combo charts

```{r}
options(scipen = 20)
ggplot(genres_df, aes(x = Genre)) +
  geom_col(aes(y = Avg_votes, fill = Genre)) +
  geom_line(aes(y = Avg_score * 10000, group = 1), color = 'red', size = 2) +
  theme(
    legend.position = 'none'
    , axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  ) +
  scale_y_continuous(sec.axis = sec_axis(~./10000, name = 'Avg_score'))

```


### Multiple plot on single panel

```{r}
ggplot(df, aes(x = IMDb.Score)) +
  geom_histogram(aes(fill = Series.or.Movie), color = 'white', alpha = 0.75, bins = 50) +
  theme(legend.position = 'none') +
  # left(rows) ~ right(columns)
  # facet_grid(.~Series.or.Movie)
  # facet_grid(Series.or.Movie ~ .)
  # facet_grid(Runtime ~ Series.or.Movie, scales = 'free')
  facet_wrap(vars(Series.or.Movie, Runtime, Trailer.Site), ncol = 5, scales = 'free')
```

```{r}
# install.packages('gridExtra')
library(gridExtra)

p1 = ggplot(df, aes(x = IMDb.Score)) +
  geom_histogram(color = 'blue', aes(fill = Series.or.Movie), alpha = 0.5) +
  labs(
    x = 'Movie rating'
    , y = 'Count of movies'
    , fill = 'Serie or movie'
    , title = 'Histogram of ratings'
  ) +
  # theme_bw() +
  # theme(legend.position = 'bottom')
  theme(legend.position = c(0.2, 0.8))

p2 = ggplot(genres_df, aes(x = Avg_score, y = Avg_votes, label = Genre)) +
  geom_point(aes(size = Freq, color = Genre)) +
  geom_label_repel() +
  theme(legend.position = 'none')

gridExtra::grid.arrange(p1, p2, ncol = 2)
```


**EXERCISE 9:** For series and movies create plot of rating from different portals (Hidden Gem, IMDb, Rotten Tomatoes, Metacritic). Hint: worth to unpivot data frame to *long* format.

```{r}
df1 = df %>% 
    pivot_longer(
    cols = ends_with("score")
    ,names_to = 'Portal'
    , values_to = 'Score'
  )

ggplot(df1, aes(x = Portal, y = Score)) +
  geom_boxplot(aes(fill = Portal)) +
  facet_wrap(vars(Portal), scales = 'free') +
  theme(legend.position = 'none')

```

**EXERCISE 10:** Which production houses are most productive and how was it changing during recent years?
```{r}

```

## Interactive data vizualization with library plotly
https://plotly.com/r/
https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

```{r}
if (!require('plotly')) install.packages('plotly')
library(plotly)
```

```{r}
data("mtcars")
```

```{r}
plot_ly(
  data = mtcars
  , x = ~wt
  , y = ~mpg
  , type = 'scatter'
  , mode = 'markers'
)

plot1 = ggplot(mtcars %>%
                 filter(wt < 5)
               , aes(x = wt, y = mpg)) +
  geom_point()

ggplotly(plot1)
```

```{r}
ggplotly(plot2)

plot_ly(
  data = df
  , x = ~IMDb.Score
  , type = 'histogram'
  # , marker = list(color = 'red', line = list(color = 'white', width = 0.5))
  , color = ~Series.or.Movie
)
```

### 3D vizualization
```{r}
model_1 = lm(mpg ~ wt + hp + as.factor(am), data = mtcars)
summary(model_1)

data_3d = expand.grid(
  wt = seq(0, 6, by = 0.1)
  , hp = seq(0, 400, by = 1)
  , am = c(0,1) %>% as.factor()
) %>%
  bind_cols(., mpg = predict(model_1, .)) %>%
  mutate(col = 'Predicted') %>%
  bind_rows(., mtcars %>% select(wt, hp, mpg, am) %>% mutate(am = as.factor(am), col = 'Real'))

plot_ly(
  data = data_3d
  , x = ~wt
  , y = ~hp
  , z = ~mpg
  , shape = ~as.factor(am)
  , color = ~col
  , size = 100
  , colors = c('orange', 'blue')
)
```

